---
title: "temporal-analysis"
author: "niwei@ccl"
date: "2021-11-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Analysis of hbo and hbr data.

```{r}
library(tidyverse)
library(data.table)
library(dplyr)
library(ggplot2)
library(readxl)
```

## load data and filter subjects

- load questionnaire data, hbo, and hbr data
- filter subjects from ASD and TD group to match IQ
- exclude 5 subjects without nirs data and 2 subjects without SPSRC data
- ASD IQ: [103,max], TD IQ: [min,137]
- final sample: 33 TD, 33 ASD

```{r}
q_total <- read_xlsx("./data/score_total.xlsx",sheet = 1,col_types = c("numeric","numeric","text","numeric","text",rep("numeric",41)))
nirs_info <- read_xlsx("./data/nirs_information.xlsx",col_types = c("text","numeric","numeric","numeric","date","numeric","text"))
load("./output/run_hbo.Rdata")
load("./output/run_hbr.Rdata")
```

```{r}
q_iq_sub <- q_total %>% 
  filter((group == 1 & (FSIQ <=143 & FSIQ >= 103)) | (group == 2 &(FSIQ <=137 & FSIQ >= 96))) %>% 
  filter(id_new != 5 & id_new != 2416 &  id_new!= 2 & id_new!= 121 & id_new!= 128 & id_new!= 23 & id_new!= 21 & id_new != 33) %>% 
  select(id_new,group,gender,name,age,FSIQ,AQ,SRS,forearm,hand,tactile,tactile_z,
         SPSRC_total,SPSRC_total_z,seeking,seeking_z,underresponse,overresponse,stability,
         tactile_over,tactile_under,tactile_seek) %>% 
  rename(id = id_new) %>% 
  as.data.table() %>% 
  setkey(id)
nirs_info <- nirs_info %>% 
  select(id,file_name) %>% 
  rename(sub = file_name) %>% 
  as.data.table() %>% 
  setkey(id)
q_iq_sub <- nirs_info[q_iq_sub] %>% 
  setkey(sub)

t.test(q_iq_sub$age[which(q_iq_sub$group==1)],q_iq_sub$age[which(q_iq_sub$group==2)])
t.test(q_iq_sub$FSIQ[which(q_iq_sub$group==1)],q_iq_sub$FSIQ[which(q_iq_sub$group==2)])
t.test(q_iq_sub$SRS[which(q_iq_sub$group==1)],q_iq_sub$SRS[which(q_iq_sub$group==2)])
t.test(q_iq_sub$AQ[which(q_iq_sub$group==1)],q_iq_sub$AQ[which(q_iq_sub$group==2)])
t.test(q_iq_sub$SPSRC_total[which(q_iq_sub$group==1)],q_iq_sub$SPSRC_total[which(q_iq_sub$group==2)])
t.test(q_iq_sub$tactile[which(q_iq_sub$group==1)],q_iq_sub$tactile[which(q_iq_sub$group==2)])
t.test(q_iq_sub$forearm[which(q_iq_sub$group==1)],q_iq_sub$forearm[which(q_iq_sub$group==2)])
t.test(q_iq_sub$hand[which(q_iq_sub$group==1)],q_iq_sub$hand[which(q_iq_sub$group==2)])

ggplot(data = q_iq_sub )+
  geom_point(aes(x = AQ, y = SPSRC_total,color = as.factor(group)))+
  scale_color_discrete(labels=c("ASD","TD"), aesthetics = c("color","fill"),type = c("#FF6666","#0099CC")) +
  theme_bw()

cor.test(q_iq_sub$AQ,q_iq_sub$SPSRC_total)
```

## HBO only

```{r}
hbo_tot <- run_hbo[q_iq_sub, on = "sub"] %>% 
  mutate(SPSRC_group = ifelse(SPSRC_total <= 403,"low", "high"),
         SPSRC_z_group = ifelse(SPSRC_total_z<=-2," <= -2 SD",ifelse(SPSRC_total_z <= -1,"-2 to -1 SD", ">= -1 SD")),
         tactile_z_group = ifelse(tactile_z<=-2," <= -2 SD",ifelse(tactile_z <= -1,"-2 to -1 SD", ">= -1 SD")),
         seeking_z_group = ifelse(seeking_z<=-2," <= -2 SD",ifelse(seeking_z <= -1,"-2 to -1 SD", ">= -1 SD")),
         tactile_group = ifelse(tactile <= 87,"low", "high"),
         tac_over_group =ifelse(tactile_over <= 29,"low", "high"),
         tac_under_group =ifelse(tactile_under <= 37,"low", "high"),
         tac_seeking_group = ifelse(tactile_seek <= 19,"low", "high"),
         seeking_group = ifelse(seeking <= 144,"low", "high"),
         over_group = ifelse(overresponse <= 134,"low", "high"),
         under_group = ifelse(underresponse <= 107,"low", "high"),
         stab_group = ifelse(stability <= 24,"low", "high"))


```

## pSTS

- compute condition difference: arm - palm
- right pSTS: channel 43 46
- left pSTS: channel 20 23

```{r}
# bilateral
hbo_sts <- hbo_tot %>% 
  group_by(id,group,SPSRC_group,tactile_group,tac_over_group,tac_under_group,tac_seeking_group,
           seeking_group,over_group,under_group,stab_group,channel,time) %>%
  summarise(con_diff = 1e06*(response[which(con == "forearm")] - response[which(con == "hand")])) %>% 
  #filter( channel == 20|channel == 23|channel == 43|channel == 46 | channel == 12| channel == 16 |channel == 39 | channel == 35) %>% 
  filter( channel == 20|channel == 23|channel == 43|channel == 46) %>% 
  mutate(time = time/3.47-2)

mean_hbo_sts <- Rmisc::summarySE(data = hbo_sts,"con_diff",groupvars = c("time","group"),na.rm = T)

ggplot(data = mean_hbo_sts,aes(x = time,y=con_diff,color = as.factor(group)))+
  geom_line()+
  #facet_wrap(~channel,scales = "free_y")+
  geom_ribbon(aes(ymax = con_diff+se,ymin = con_diff-se,fill = as.factor(group)),alpha = 0.25,linetype=0)+
  scale_color_discrete(labels=c("ASD","TD"), aesthetics = c("color","fill"),type = c("#FF6666","#0099CC")) +
  scale_x_continuous(breaks = c(-2,0,2,4,6,8,10,12,14,16))+
  scale_y_continuous(breaks = seq(-0.1,0.1,0.02))+
  #geom_rect(aes(xmax = 40/3.47-5,xmin = 33/3.47-5,ymax = -0.088,ymin = -0.09),color = "yellow",linetype=0) +
  geom_vline(aes(xintercept = 0))+
  geom_vline(aes(xintercept = 6))+
  labs(title = "Bilateral pSTS activation to arm vs palm",color = "Group",fill = "Group")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# right

hbo_sts <- hbo_tot %>% 
  group_by(id,group,SPSRC_group,tactile_group,tac_over_group,tac_under_group,tac_seeking_group,
           seeking_group,over_group,under_group,stab_group,channel,time) %>%
  summarise(con_diff = 1e06*(response[which(con == "forearm")] - response[which(con == "hand")])) %>% 
  #filter( channel == 20|channel == 23|channel == 43|channel == 46 | channel == 12| channel == 16 |channel == 39 | channel == 35) %>% 
  filter( channel == 43|channel == 46) %>% 
  mutate(time = time/3.47-2)

mean_hbo_sts <- Rmisc::summarySE(data = hbo_sts,"con_diff",groupvars = c("time","group"),na.rm = T)

ggplot(data = mean_hbo_sts,aes(x = time,y=con_diff,color = as.factor(group)))+
  geom_line()+
  #facet_wrap(~channel,scales = "free_y")+
  geom_ribbon(aes(ymax = con_diff+se,ymin = con_diff-se,fill = as.factor(group)),alpha = 0.25,linetype=0)+
  scale_color_discrete(labels=c("ASD","TD"), aesthetics = c("color","fill"),type = c("#FF6666","#0099CC")) +
  scale_x_continuous(breaks = c(-2,0,2,4,6,8,10,12,14,16))+
  scale_y_continuous(breaks = seq(-0.1,0.1,0.02))+
  #geom_rect(aes(xmax = 40/3.47-5,xmin = 33/3.47-5,ymax = -0.088,ymin = -0.09),color = "yellow",linetype=0) +
  geom_vline(aes(xintercept = 0))+
  geom_vline(aes(xintercept = 6))+
  labs(title = "Right pSTS activation to arm vs palm",color = "Group",fill = "Group")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))


# left

hbo_sts <- hbo_tot %>% 
  group_by(id,group,SPSRC_group,tactile_group,tac_over_group,tac_under_group,tac_seeking_group,
           seeking_group,over_group,under_group,stab_group,channel,time) %>%
  summarise(con_diff = 1e06*(response[which(con == "forearm")] - response[which(con == "hand")])) %>% 
  #filter( channel == 20|channel == 23|channel == 43|channel == 46 | channel == 12| channel == 16 |channel == 39 | channel == 35) %>% 
  filter( channel == 20|channel == 23) %>% 
  mutate(time = time/3.47-2)

mean_hbo_sts <- Rmisc::summarySE(data = hbo_sts,"con_diff",groupvars = c("time","group"),na.rm = T)

ggplot(data = mean_hbo_sts,aes(x = time,y=con_diff,color = as.factor(group)))+
  geom_line()+
  #facet_wrap(~channel,scales = "free_y")+
  geom_ribbon(aes(ymax = con_diff+se,ymin = con_diff-se,fill = as.factor(group)),alpha = 0.25,linetype=0)+
  scale_color_discrete(labels=c("ASD","TD"), aesthetics = c("color","fill"),type = c("#FF6666","#0099CC")) +
  scale_x_continuous(breaks = c(-2,0,2,4,6,8,10,12,14,16))+
  scale_y_continuous(breaks = seq(-0.1,0.1,0.02))+
  #geom_rect(aes(xmax = 40/3.47-5,xmin = 33/3.47-5,ymax = -0.088,ymin = -0.09),color = "yellow",linetype=0) +
  geom_vline(aes(xintercept = 0))+
  geom_vline(aes(xintercept = 6))+
  labs(title = "Left pSTS activation to arm vs palm",color = "Group",fill = "Group")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

```

## Somatosensory cortex

- left somatosensory cortex: channel 13 14 17 18 21

```{r}
hbo_sm <- hbo_tot %>% 
  filter(channel == 13|channel == 17|channel == 14|channel ==18|channel == 21) %>% 
  mutate(time = time/3.47-2,response = response*1e06)

# group comparison
mean_hbo_sm <- Rmisc::summarySE(data = hbo_sm,"response",groupvars = c("time","group"),na.rm = T)
ggplot(data = mean_hbo_sm,aes(x = time,y=response,color = as.factor(group)))+
  geom_line()+ 
  geom_ribbon(aes(ymax = response+se,ymin = response-se,fill = as.factor(group)),alpha = 0.25,linetype=0)+
  scale_color_discrete(labels=c("ASD","TD"), aesthetics = c("color","fill"),type = c("#FF6666","#0099CC")) +
  scale_x_continuous(breaks = c(-2,0,2,4,6,8,10,12,14,16))+
  scale_y_continuous(breaks = seq(-0.1,0.1,0.02))+
  geom_vline(aes(xintercept = 0))+
  geom_vline(aes(xintercept = 6))+
  labs(title = "Left somatosensory cortex activation under both condition",color = "Group",fill = "Group")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# tactile seeking group
mean_hbo_sm <- Rmisc::summarySE(data = hbo_sm,"response",groupvars = c("time","tac_seeking_group"),na.rm = T)
ggplot(data = mean_hbo_sm,aes(x = time,y=response,color = as.factor(tac_seeking_group)))+
  geom_line()+ 
  #facet_wrap(~channel,scales = "free_y")+
  geom_ribbon(aes(ymax = response+se,ymin = response-se,fill = as.factor(tac_seeking_group)),alpha = 0.25,linetype=0)+
  scale_color_discrete(labels=c("Lower","Higher"), aesthetics = c("color","fill"),type =c("#6E7783","#77AAAD")) +
  scale_x_continuous(breaks = c(-2,0,2,4,6,8,10,12,14,16))+
  scale_y_continuous(breaks = seq(-0.1,0.1,0.02))+
  geom_vline(aes(xintercept = 0))+
  geom_vline(aes(xintercept = 6))+
  labs(title = "Left somatosensory cortex activation under both condition",color = "Tactile seeking",fill = "Tactile seeking")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```
## HBO and HBR

- pSTS

```{r}

```